# ---- Week 5 Seminar Data ----

# ---- Load Libraries ----

```{r}
library(ggplot2)
library(tidyverse)
library(tidyr)
library(naniar)
```

# ----- Upload Data File -----

```{r}
strengthdata <- readxl::read_excel("/Users/seanmccrone/Desktop/MASTERS DEGREE/Course Material/B1700/Week 5/Seminar Data/Strength Testing Data.xlsx")
```

# ----- Summary of Data File -----

```{r}
summary(strengthdata)
```

# ----- Cleaning the Data File -----

##### Converting Variables from Character to Numeric/Integer

```{r}
strengthdata <- strengthdata %>%
  mutate(across(2:5, ~as.integer(.x)))
strengthdata
```

##### Code to remove 'NA' values, now that we have converted to numeric/integer

```{r}
strengthdata <- na.omit(strengthdata)
```

##### Code to rename the variables

```{r}
strengthdata <- strengthdata %>%
  rename(PlayerName = `Player Name`,
         HamForL = `Hamstring Force (L)`,
         HamForR = `Hamstring Force (R )`,
         GroForL = `Groin Strength (L)`,
         GroForR = `Groin Strength (R )`)
```

##### Code for summary of the strengthdata

```{r}
summary(strengthdata)
```

# ---- Creating Visuals ----

##### Code to stack all variables representing force on top of each other per player

```{r}
strengthdata_long <- strengthdata %>%
  pivot_longer(cols = c(HamForL, HamForR, GroForL, GroForR), names_to = "Variable", values_to = "Force")

ggplot(strengthdata_long, aes(fill = Variable, x = PlayerName, y = Force)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Stacked Bar Plot", x = "Player Name", y = "Force")
```

##### This code creates a grouped bar graph, showing the 4 variables per player side by side

```{r}
strengthdata_long <- strengthdata %>% 
  pivot_longer(cols = c(HamForL, HamForR, GroForL, GroForR), 
               names_to = "Variable", 
               values_to = "Force")

ggplot(strengthdata_long, aes(x = PlayerName, y = Force, fill = Variable)) + 
  geom_bar(stat = "identity", position = position_dodge(), width = 0.5) +
  labs(title = "Grouped Bar Graph", x = "Player Name", y = "Force")
```

##### This code removes the outliers found in the dataset 'strengthdata'

```{r}
remove_outliers <- function(df, columns, z_threshold = 2) {
  df_cleaned <- df
  for (col in columns) {
    z_scores <- scale(df_cleaned[[col]])
    outliers <- abs(z_scores) > z_threshold
    df_cleaned <- df_cleaned[!outliers, ]
  }
  df_cleaned
}
strengthdata2 <- remove_outliers(strengthdata, c("HamForL", "HamForR","GroForL", "GroForR"))
```
